% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
]{article}
\usepackage{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\newcounter{none} % for unnumbered tables
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

\section{The Gauntlet that Catches Itself: Pre-Registered AI Evaluation
Infrastructure with In-Production Bar-Weakness Detection, with a First
PASS Mechanism
Characterization}\label{the-gauntlet-that-catches-itself-pre-registered-ai-evaluation-infrastructure-with-in-production-bar-weakness-detection-with-a-first-pass-mechanism-characterization}

\textbf{Author:} Alexander Rodabaugh (Fathom Lab) \textbf{Date:}
2026-05-27 (revised 2026-05-28, v6 --- adds §14 documenting Layers 5 and
6 of the instrumented recursion frame) \textbf{Substrate:} styxx 7.7.10
· nineteen reference baselines on the dark-core benchmark · git origin
\texttt{fathom-lab/styxx@main} \textbf{Status:} preprint, in-session
synthesis

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Abstract}\label{abstract}

We present a publicly-reproducible pre-registered AI evaluation gauntlet
for hallucination/misconception detection, instantiated against a
108-record benchmark of cross-vendor consensus errors. The contribution
has three parts: (1) a meta-property of the infrastructure --- the
gauntlet \textbf{caught two of its own bar weaknesses in production
within a single session}, replaced them with regression-tested controls,
and re-scored all existing submissions honestly under the strengthened
bars; (2) the first method to PASS the strengthened bars ---
\textbf{gpt-4o-mini in critique mode}, AUC 0.95, achieved at pre-stated
28\% probability with all six AUC-range predictions inside their
pre-stated ranges; and (3) a published-then-falsified-then-corrected
mechanism measurement: the v1 FINDING attributed the PASS to a 91\%
within-model generation-vs-critique asymmetry; a subsequent in-session
demo revealed that cosine-similarity proxies conflate topical relevance
with truth-value agreement; a directional NLI re-test (v2) put strict
within-model asymmetry at 5.88\% but suffered an 85\% UNCLEAR rate; a
third measurement (v3) with forced single-character T/F/U output
resolved the UNCLEAR artifact entirely (0\% on dark-core, 13\% on
TruthfulQA) and landed the corrected rates inside pre-stated bands:
\textbf{5.88\% on dark-core and 17.00\% on TruthfulQA}. Three iterations
of measurement, each pre-registered, each in git history. We record
\textbf{sixteen in-session falsifications + four resolutions} including
two of our own published FINDING's central methodology, the corrected v3
measurement, the same-session self-falsification of v4's own
forward-looking claim about \texttt{styxx.critique\_detector}
(documented in §13, closed in commit \texttt{0e97598}), and --- newly
added in v6 --- two instrumented frames of the same discipline
(documented in §14): a substrate-grounded auditor
(\texttt{styxx.agent\_audit}, Layer 5) that verified 13/13
pre-registered session-output claims against the substrate, and the
paper's own published primitive (\texttt{styxx.critique\_detector},
Layer 6) applied to 13 TRUE paraphrases plus 5 deliberate FALSE controls
drawn from the paper's own §11.5 and §13 source passages, with both
pre-stated kill-gates un-fired (18/18 PASS at threshold-0.50). The
corrected mechanism description for the gauntlet PASS is
\textbf{out-of-context critique}: gpt-4o-mini reliably rejects
labeled-misconception text presented as a candidate, regardless of
whether it would have generated that text itself. We argue that the moat
of disciplined AI evaluation is the \textbf{recursive pattern}: bars
revise under empirical pressure; the discipline of pre-registration
BEFORE each run makes wrongness \emph{visible}; the infrastructure
improves itself on a session timescale; when published FINDINGs are
themselves wrong, the discipline catches and revises them; when
methodology iterations are needed to land a clean measurement, the
discipline pre-registers each iteration honestly; when the paper itself
makes forward-looking claims about its own substrate, the same
discipline catches and closes the gap; AND when those claim-vs-substrate
checks can be \textbf{instrumented and pre-registered against} --- both
via a custom-built session-output auditor and via the paper's own
published primitive --- the recursion produces falsifiable kill-gates on
its own results, run end-to-end on a session timescale. All artifacts
(benchmark, gauntlet code, nineteen baseline submissions, ten FINDING
documents, sixteen+four pre-stated-then-published predictions, the first
PASS event, the asymmetry v1/v2/v3 measurements, the §13 self-audit
closure commit, the §14 instrumented-recursion-frame artifacts including
18 cross-model context-grounded faithfulness scores) are reconstructible
from public git history at commit-level granularity.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{1. Introduction}\label{introduction}

The AI evaluation literature has a well-documented credibility crisis:
benchmarks leak into training data; bars get gamed by surface-form
features; closed-vendor evaluation pipelines cannot be reproduced
independently; reported state-of-the-art numbers fail to replicate on
follow-up corpora.\footnote{e.g., Hendrycks et al., 2020; Ribeiro et
  al., 2020; Recht et al., 2019. The ``evaluation crisis'' literature is
  too large to cite exhaustively.} The proposed responses are familiar
--- held-out test sets, contamination audits, formal pre-registration
--- but each has its own failure modes. Pre-registration of a fixed bar
set, in particular, has a flaw the empirical-evaluation literature has
not adequately addressed: \textbf{what happens when the pre-registered
bars themselves turn out to be gameable?}

This paper documents a real-world instance where exactly that happened,
and the corrective mechanism that followed. On 2026-05-27, the styxx
project shipped a public-challenge runner (``the gauntlet'') with two
detection bars (D1, D2). Within hours, an external-style sanity-check
submission --- a 30-line token-overlap heuristic --- accidentally PASSed
both bars. Investigation revealed a length-confound in the benchmark's
\texttt{expected\_consensus} field that no design-time review had
surfaced. The fix shipped six minutes later as a third bar (D3
length-control), with a regression test that the length-only oracle must
fail by construction. A subsequent \textbf{systematic confound audit},
pre-registered with eight feature predictions committed to public git
BEFORE running, revealed a second orthogonal artifact
(capitalized-token-ratio, inverted) that became the D4 bar. Three
pre-registered detection submissions tested under the strengthened v3
bars all failed, in their predicted modal outcome regions. The empirical
floor compounds.

What we record here is not ``we found the truth about hallucination
detection.'' We did not. The dark core stays dark to every method we
tested. The contribution is the \textbf{infrastructure pattern}: how to
build evaluation gauntlets that \emph{catch their own validity
weaknesses in production}, and what the costs and shape of that
discipline look like when run rigorously.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{2. The empirical floor: the seven-method dark-core
arc}\label{the-empirical-floor-the-seven-method-dark-core-arc}

The benchmark for this paper is
\texttt{darkcore\_benchmark\_2026\_05\_27.json}, n=108 records, four
classes (truth=55, folklore=34, factual-error=13, pseudoscience=6). Each
record contains a question, an \texttt{expected\_consensus} (the
consensus answer produced by a three-vendor council: gpt-4o-mini +
Qwen2.5-3B + gemma-2-2b-it), and a hand-curated class label. The
benchmark was constructed for the \emph{Decorrelation Ceiling} paper
\footnote{\texttt{papers/PAPER\_decorrelation\_ceiling\_2026\_05\_27.md}
  in the same repository, also released 2026-05-27.}, whose central
claim is that \textbf{reference-free divergence methods detect
cross-vendor errors iff a decorrelated competing representation is
available} --- i.e., the methods fail on the subset where all three
vendors converge on the same wrong answer (``the dark core'').

The Decorrelation Ceiling paper tested seven methods (the ``seven-method
floor''):

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Dark Matter} --- perturbation-fragility under prompt
  rephrasing (closed-negative on the dark core)
\item
  \textbf{CVPD} --- cross-vendor pairwise disagreement (closed-negative)
\item
  \textbf{JD} --- justification-divergence under explanation prompts
  (closed-negative)
\item
  \textbf{ICT} --- neutral injection of contrary content
  (closed-negative)
\item
  \textbf{ICT-folklore} --- folklore-prime variant (28/30
  already-corrected shortfall)
\item
  \textbf{ICT-authoritative} --- authoritative-prime variant (same
  shortfall shape)
\item
  \textbf{Dark-core classifier} --- sentence-transformer + balanced
  one-vs-rest LR (K1=0.42, K3=0.36 → fails 2/3 classification bars)
\end{enumerate}

The seven-method floor is the empirical ground state: each is a method
we ran honestly, each is pre-registered with bars committed BEFORE data,
each is closed-negative on the partition that matters. The gauntlet's
contribution to the empirical landscape is not its own methods --- it's
the public-challenge runner that lets \emph{anyone else} try.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{3. The gauntlet
infrastructure}\label{the-gauntlet-infrastructure}

\texttt{styxx.gauntlet} (shipped 7.7.5, hardened 7.7.8 + 7.7.9) is a
public-challenge runner with two task modes:

\begin{itemize}
\tightlist
\item
  \textbf{Classification:} the user's method takes a question, returns a
  predicted class label. Bars: K1 (folklore F1 in-distribution ≥ 0.70),
  K2 (4-way accuracy ≥ 0.65), K3 (cross-corpus folklore F1 ≥ 0.60,
  load-bearing).
\item
  \textbf{Detection:} the user's method takes (question, response),
  returns a misconception score. Bars: D1 (misconception vs truth AUC ≥
  0.70), D2 (folklore vs truth AUC ≥ 0.70), D3 (length-control delta ≥
  0.10, added 7.7.8), D4 (capitalization-control delta ≥ 0.10, added
  7.7.9).
\end{itemize}

The interface is deliberately framework-agnostic: pass a Python callable
(\texttt{module:attr} spec), the runner imports and applies it to every
record in the benchmark, scores it against pre-registered bars, and
returns a structured \texttt{GauntletResult}. PASS is the conjunction
over all bars.

The benchmark, the bars, and the runner all ship as installed package
data (\texttt{pip\ install\ styxx}), making the gauntlet
\textbf{publicly reproducible without git access or remote benchmark
downloads}.

The four phases of a submission:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Submitter writes a Python callable conforming to the task signature.
\item
  Submitter runs the gauntlet locally; receives a structured JSON
  result.
\item
  Submitter opens a PR adding a row to \texttt{LEADERBOARD.md} with the
  submission JSON attached.
\item
  CI re-runs the gauntlet on the submitter's method against the locked
  benchmark; if scores match, the PR is mergeable.
\end{enumerate}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{4. The pre-registration
discipline}\label{the-pre-registration-discipline}

For every detection submission tested in this paper, we follow a
five-step discipline:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Write a pre-stated prediction document}
  (\texttt{PRE\_STATED\_PREDICTION.md}) committing to expected AUC
  ranges, outcome probability distribution, and direction-of-effect
  hypotheses.
\item
  \textbf{Commit the prediction + the method file (un-run) to public
  origin} before any gauntlet invocation. The git commit hash is the
  prereg witness.
\item
  \textbf{Run the gauntlet exactly once.} No re-running on the same
  submission. No hyperparameter sweeps. No ``trying a different model.''
\item
  \textbf{Publish the result to the public LEADERBOARD regardless of
  outcome.} Modal-prediction validation, AUC-range hits, and
  direction-of-effect misses are all recorded honestly.
\item
  \textbf{If a submission surfaces a bar weakness, the bar gets revised}
  (with a regression test) and ALL prior submissions get re-scored under
  the new bars. Original-bar scores are preserved alongside the new
  ones; nothing is hidden.
\end{enumerate}

The discipline is asymmetric: it constrains us (the project) more than
it constrains submitters. A submitter can run their method many times
privately, only publishing the best result; we publish every gauntlet
run we make against ourselves. The asymmetry is intentional. The
seven-method floor is \emph{our} floor; future submissions either beat
the floor (the synthesis revises) or fail (the floor compounds across
submissions).

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{5. The bars-catching-themselves pattern: D3 →
D4}\label{the-bars-catching-themselves-pattern-d3-d4}

The central empirical demonstration of this paper is two consecutive bar
revisions that emerged from operating the gauntlet, not from designing
it.

\subsubsection{5.1. D3 --- caught by
accident}\label{d3-caught-by-accident}

A 30-line token-overlap detector (``Baseline-007''):
\texttt{score\ =\ (1\ −\ hedge\_density)\ ×\ (1\ −\ novelty\_density)},
no model, no training data. Submitted as a sanity-check to populate the
detection side of the leaderboard. Expected outcome: noise-level
baseline.

Actual outcome: \textbf{PASS=true, D1=0.864, D2=0.922, 2/2 bars
cleared.} The first-ever PASS on the gauntlet, from a method that should
not have passed.

Investigation revealed a length-confound in the benchmark's
\texttt{expected\_consensus} field. The class-conditional mean response
length is sharply discriminative:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}ll@{}}
\toprule\noalign{}
class & mean \texttt{expected\_consensus} length (words) \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
truth & 3.9 \\
factual-error & 6.6 \\
pseudoscience & 6.5 \\
folklore & 7.5 \\
\end{longtable}
}

A pure length-only oracle (\texttt{score\ =\ len(response.split())}) was
run against the same partitions and scored AUC ≈ 0.79--0.80 on both D1
and D2. The v1 bars (AUC ≥ 0.70) were trivially gameable by detectors
whose score positively correlates with response length.

Six minutes later, styxx 7.7.8 shipped with the D3 length-control bar:
detector AUC must beat the length-only oracle's AUC by ≥ 0.10 on BOTH
partitions. A regression test
(\texttt{test\_length\_oracle\_passes\_D1\_D2\_but\_fails\_D3}) was
added the same commit: the length oracle (whose score IS length)
trivially passes D1+D2 by construction but D3 must fail by construction
(delta = 0). Baseline-007 was re-scored: now 2/3, NOT a PASS. The
original v1 score is preserved alongside the v2 score in the submission
JSON.

This took six minutes from PASS to patch.

\subsubsection{5.2. D4 --- caught by deliberate
scan}\label{d4-caught-by-deliberate-scan}

After D3, the next disciplined question: \textbf{what other surface
features game the bars?} We defined eight candidate oracle-detectors:

\begin{itemize}
\tightlist
\item
  \texttt{word\_length} (calibration --- the existing D3 oracle)
\item
  \texttt{char\_length}
\item
  \texttt{sentence\_count}
\item
  \texttt{question\_mark\_count}
\item
  \texttt{exclamation\_count}
\item
  \texttt{capitalized\_token\_ratio}
\item
  \texttt{hedge\_density}
\item
  \texttt{type\_token\_ratio}
\end{itemize}

For each, we computed: D1 AUC, D2 AUC, Spearman ρ to word\_length
(orthogonality measure). An oracle that passes a bar AT ρ \textless{}
0.5 with word\_length is a \emph{candidate orthogonal confound} --- a
feature that games the bars without being a length proxy.

Pre-stated predictions for all 8 features × 2 partitions = 16 AUC ranges
were committed to public origin BEFORE running. The audit then revealed
\texttt{capitalized\_token\_ratio} as a genuine orthogonal confound ---
but \textbf{inverted from the predicted direction}:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}lllll@{}}
\toprule\noalign{}
oracle & D1 AUC raw & D1 AUC abs & direction & ρ → length \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\texttt{capitalized\_token\_ratio} & 0.296 & \textbf{0.704} & inverted &
−0.343 \\
\end{longtable}
}

The mechanism: truth responses are canonical short answers like
``Paris'', ``Newton'', ``1789'' --- strings where the capitalized-token
ratio is structurally near 1.0. Folklore restatements are full sentences
where lowercase function words dilute the proper-noun density. Truth has
\emph{higher} cap-ratio, not lower.

The pre-stated audit code missed this initially: my flagging logic
checked \texttt{AUC\ ≥\ d1\_bar}, the positive direction only. The
cap-ratio's raw AUC was 0.30, below the 0.70 threshold, so the first
audit reported \texttt{n\_orthogonal\_confounds\_found:\ 0}. \textbf{My
code was confounded by the same direction-blindness as my prediction.}
Fixed by adding \texttt{D1\_AUC\_abs\ =\ max(auc,\ 1−auc)} and flagging
based on absolute AUC. Re-ran; cap-ratio surfaced as a candidate
orthogonal confound.

styxx 7.7.9 shipped with D4 (capitalization-control delta ≥ 0.10) plus a
regression test
(\texttt{test\_capratio\_oracle\_passes\_D1\_D2\_but\_fails\_D4}). PASS
now requires D1 ∧ D2 ∧ D3 ∧ D4.

\subsubsection{5.3. The recursion is the
property}\label{the-recursion-is-the-property}

The chain extends:

\begin{itemize}
\tightlist
\item
  \textbf{D3} discovered by \emph{accident} (an external-style sanity
  submission unexpectedly PASSed)
\item
  \textbf{D4} discovered by \emph{deliberate scan} (the audit primitive
  runs on demand)
\end{itemize}

Future confounds would be caught by the same pattern. The audit
primitive \texttt{styxx.gauntlet.audit\_confounds()} is part of the
public API; anyone running \texttt{styxx\ gauntlet-audit-confounds} gets
the per-feature AUC + ρ-to-length table. Operator-territory follow-ups
already identified: \texttt{numeric\_token\_ratio},
\texttt{single\_token\_flag}, \texttt{uppercase\_ratio} --- none
implemented yet but cheap to add.

The bars compound: a method that clears all four has demonstrated signal
beyond the length artifact AND the cap-ratio artifact AND scored AUC ≥
0.70 on both broad and folklore axes. Each new D-bar adds a
confound-rejection commitment that future detectors must satisfy.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{6. Thirteen in-session
falsifications}\label{thirteen-in-session-falsifications}

We name what was wrong, in public, in git history. The thirteen
in-session falsifications of the 2026-05-27 arc:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{C1-profile ≤ 0.20 register-law bar} (Pareto-frontier finding):
  C10 deliberately written in the law's voice scored composite 0.264,
  missing the bar. The Pareto finding was revised in place.
\item
  \textbf{\texttt{set\_session} doesn't propagate} (product-exploration
  finding): falsified by per-agent routing --- \texttt{set\_session}
  DOES propagate via \texttt{STYXX\_AGENT\_NAME}.
\item
  \textbf{ICT-folklore auto-verdict PASS} (probe label bug, commit
  \texttt{cc3435c}).
\item
  \textbf{ICT-authoritative auto-verdict PASS} (same label bug shape,
  commit \texttt{a6d7a7e}).
\item
  \textbf{styxx 7.7.5 wheel-bundling miss} (benchmark JSON not included
  in installed wheel).
\item
  \textbf{The gauntlet's v1 detection bars being length-gameable} (D3
  discovery).
\item
  \textbf{The cap-ratio orthogonal confound} (D4 discovery): predicted
  positive direction; actual inverted.
\item
  \textbf{Baseline-010's NLI direction}: predicted folklore entails
  question; actual reversed.
\item
  \textbf{Baseline-011's magnitudes underpredicted}: D1=0.811 vs
  predicted 0.55-0.72; D2=0.897 vs predicted 0.58-0.78. Outcome band
  correct.
\item
  \textbf{Baseline-012 ``scaling solves the signal''}: predicted 60\%
  modal 3/4 at gpt2-large; actual 1/4 --- scaling within gpt2 family
  DEGRADES signal.
\item
  \textbf{Baseline-018 dual-LM composite}: predicted 22\% PASS via ``use
  the scaling curve as a feature''; actual 2/4 --- composite signal
  weaker than single-LM at both endpoints.
\item
  \textbf{Asymmetry §11 prevalence underpredicted}: predicted 50-80\% HH
  quadrant; actual 91.18\% --- well above pre-stated upper bound (second
  magnitude-underprediction in the arc).
\item
  \textbf{(Implicit) the ``first PASS requires cross-vendor or
  fine-tuned classifier'' framing in the v2 of this paper}: §11
  establishes that the first PASS came from a same-vendor model in
  critique mode, not from a structurally new vendor. The expectation
  that operator-territory resources were strictly needed was wrong; what
  was needed was a \emph{prompting-mode shift}.
\end{enumerate}

The pattern across (7), (8), (9), (10), (12) sharpens into a durable
domain-specific calibration lesson:

\begin{itemize}
\tightlist
\item
  \textbf{Direction-of-effect predictions} on this domain are
  systematically the worst --- (7), (8) both miss direction.
\item
  \textbf{Magnitude predictions} are roughly two-thirds reliable on the
  lower end but \textbf{systematically too conservative on the upper
  end} --- (9), (12) both underpredict.
\end{itemize}

Calibration record across pre-stated artifacts this session:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
pre-stated artifact
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
predictions
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
inside-range count
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
outcome-band call
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
direction call
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Baseline-006 (char TF-IDF) & binary tie & exact match & confirmed &
n/a \\
Baseline-008 (embedding similarity) & 5 ranges + 4 bands & 5/5 + 60\%
band & well-calibrated & confirmed \\
Baseline-009 (residualized) & 6 ranges + 5 bands & 4/6 + 30\% band &
well-calibrated & confirmed \\
Baseline-010 (NLI) & 6 ranges + 6 bands & 0/6 (direction wrong) + 20\%
band & band-correct & falsified \\
Baseline-011 (gpt2-124M) & 6 ranges + 6 bands & 4/6 (magnitudes
underpredicted) + 15\% band & band-correct & confirmed \\
Baselines 012-017 (scaling sweep) & per-baseline ranges & gradually
improving as anchors bracket & various & all confirmed \\
Baseline-018 (scaling-residual) & 6 ranges + 6 bands & partial inside +
25\% band & band-correct & confirmed \\
\textbf{Baseline-019 (FIRST PASS)} & \textbf{6 ranges + 6 bands} &
\textbf{6/6 inside + 28\% PASS band} & \textbf{HIT} & \textbf{confirmed
(9th)} \\
Asymmetry §11 & 4 quadrant ranges & 1/4 inside (HH underpredicted) &
strong-effect direction & n/a \\
Confound audit & 16 ranges + 6 joint & 3/16 + 3 of 6 joint falsified &
poorly calibrated & partial \\
\end{longtable}
}

\textbf{The session's discipline lesson, in one line:}
\emph{direction-of-effect predictions on RLHF behavior are unreliable;
magnitude upper bounds are systematically too conservative when the
mechanism is real. Future predictions should widen upper tails on both
AUC and prevalence ranges.}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{7. The detection frontier under v3
bars}\label{the-detection-frontier-under-v3-bars}

After D3 + D4, the gauntlet has documented receipts for what does NOT
clear the dark core:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2000}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
method class
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
best submission
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
bars passed
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
failed bar
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
reason
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Surface lexical & Baseline-007 (token overlap) & 3/4 & D3 & tracks
length too closely \\
Classical NLP & Baseline-005/006 (TF-IDF) & 1/3 on classification & K1,
K3 & folklore class lacks lexical-form signature \\
Raw semantic embedding & Baseline-008 (sent-T cosine) & 3/4 & D3 &
structurally equivalent to length on D1 \\
Length-corrected embedding & Baseline-009 (residualized) & 1/4 & D1, D3,
D4 & residualization subtracts signal rather than adding \\
Pre-trained NLI cross-encoder & Baseline-010 (R-as-premise entailment) &
0/4 & all four & MNLI training doesn't transfer to
factual-restatement \\
\end{longtable}
}

That is substantial coverage of the ``cheap surface-text'' method space.
The bars are now forcing the research budget toward more substantive
features. Methods that have not yet been tested:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Cross-vendor consensus disagreement} --- score by inter-vendor
  disagreement after re-eliciting the answer from ≥ 2 vendors. Requires
  vendor keys; estimated PASS probability 25--40\%.
\item
  \textbf{Perplexity against a calibrated prior LM} --- score response
  surprise given the question, conditioned on a reference distribution.
  Requires a small open LM; uncertain expected signal.
\item
  \textbf{Knowledge-graph lookup} --- query Wikidata / DBPedia, score
  response-KG agreement. Operator-territory; multi-hour build.
\item
  \textbf{Fine-tuned classifier on (q, r) pairs} --- retrain
  Baseline-002's architecture jointly on paired inputs. Risks length
  leakage at training time.
\end{enumerate}

The remaining frontier requires investment beyond surface-text
manipulation. That is the discipline pattern's contribution: it forces
the \emph{next} bet to be a more substantive one.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{8. How the discipline
scales}\label{how-the-discipline-scales}

The pattern documented in this paper has properties that we argue
generalize beyond this benchmark:

\textbf{Falsifiability is a substrate, not a step.} Every artifact in
this arc has a pre-stated commit hash. The eight in-session
falsifications are visible because they were pre-stated. The
bars-catching-themselves recursion is visible because the original
bar-set was committed before the submission that exposed its weakness.
Without pre-registration, the same eight events would have looked like
development noise.

\textbf{Bars revise under empirical pressure, not under design
pressure.} D3 was not foreseen by anyone reviewing the v1 bars; it was
foreseen by a sanity-check submission that exposed the artifact in
production. D4 was foreseen by deliberate scan AFTER D3 existed. The
bars cannot be made artifact-free at design time; they get hardened by
being used.

\textbf{The discipline asymmetry is the moat.} Submitters can run their
methods privately many times before submitting their best result. We
must publish every gauntlet run we make against ourselves, every
prediction we lock, every falsification that follows. The asymmetry is
intentional: it makes the floor's credibility cost more for us than the
wins it gives to submitters. Over time, this compounds --- the floor
accumulates rigor faster than any single submission can erode.

\textbf{Calibration improves where the pattern catches misses.} The
direction-of-effect lesson from cap-ratio and NLI is now durable. Future
pre-stated AUC predictions on this domain will include direction as a
separate sub-prediction. The discipline produces domain-specific
calibration improvements that no design-time review could have
predicted.

\textbf{Reproducibility is the byproduct of the pattern, not a separate
property.} Every artifact in this paper has a public git commit, a
public PyPI release-target (7.7.9, ready for publish), a public
benchmark JSON, a public LEADERBOARD row, and a structured submission
JSON. \texttt{pip\ install\ styxx==7.7.9} +
\texttt{styxx\ gauntlet\ -\/-method\ \textless{}spec\textgreater{}\ -\/-task\ \textless{}task\textgreater{}}
reproduces every result on the leaderboard.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{9. Honest limitations}\label{honest-limitations}

We name what this paper does NOT establish:

\begin{itemize}
\tightlist
\item
  \textbf{The dark core is not solved.} No method tested in this arc has
  cleared D3+D4. The seven-method floor + the gauntlet's v3 bars stand.
  Future submissions may pass; until then, the empirical claim is ``the
  methods we have tested fail.''
\item
  \textbf{The benchmark is single-vendor on the labels.} The
  three-vendor council is gpt-4o-mini + Qwen2.5-3B + gemma-2-2b-it.
  Cross-vendor generalization to other major vendors (Anthropic,
  Google's Gemini-2.5, Mistral) is untested.
\item
  \textbf{The benchmark is English-language with Western cultural
  priors.} Cross-language / cross-cultural generalization is a known
  open follow-up.
\item
  \textbf{The D3 and D4 thresholds (0.10) are judgment calls.} They were
  pre-registered before submission outcomes, but they could be wrong. If
  serious submissions cluster at D3-delta ∈ {[}0.05, 0.10{]} without
  obviously gaming the artifact, the threshold may need revision based
  on the submission distribution --- but only based on submission data,
  never to make the bar easier.
\item
  \textbf{\texttt{audit\_confounds} audits only the 8 features we
  defined.} Additional candidate confounds (numeric-token-ratio,
  single-token-flag, uppercase-ratio) are not implemented. Future
  submissions may surface new orthogonal artifacts that the current
  audit cannot detect.
\item
  \textbf{N = 108 is small.} AUC estimates have meaningful confidence
  intervals at this corpus size. The bars (≥ 0.70) are well above chance
  but not above plausible variance for cherry-picked submissions; the
  discipline pattern compensates by requiring pre-registration
  \emph{before} the submission sees the corpus.
\end{itemize}

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{10. The first PASS event: gpt-4o-mini critique
mode}\label{the-first-pass-event-gpt-4o-mini-critique-mode}

After 18 pre-registered detection baselines all FAILed v3 bars ---
including a full LM-likelihood scaling sweep across two model families
(Baselines 011-017) and a composite scaling-residual detector
(Baseline-018) --- \textbf{Baseline-019} broke through with 4/4 bars at
a pre-stated 28\% PASS probability that landed cleanly.

\subsubsection{The method}\label{the-method}

Score \texttt{=\ P("NO"\ \textbar{}\ critique\ prompt)} where the prompt
is:

\begin{verbatim}
Question: {q}

Proposed answer: {expected_consensus}

Is the proposed answer factually correct? Respond with only the single word YES or NO.
\end{verbatim}

extracted from the first-token logprobs of \texttt{gpt-4o-mini} at
temperature 0.

\subsubsection{The result}\label{the-result}

\begin{verbatim}
D1 misconception AUC       = 0.8938 ≥ 0.70 ✓
D2 folklore AUC            = 0.9548 ≥ 0.70 ✓
D1 − length oracle AUC     = 0.1041 ≥ 0.10 ✓   ← the lone blocker, finally cleared
D2 − length oracle AUC     = 0.1505 ≥ 0.10 ✓
D1 − cap-ratio oracle AUC  = 0.1902 ≥ 0.10 ✓
D2 − cap-ratio oracle AUC  = 0.1628 ≥ 0.10 ✓
──────────────────────────────────────────────
Bars passed                = 4 / 4  →  PASS
\end{verbatim}

All six AUC-range predictions held inside the pre-stated ranges.
Direction confirmed for the ninth consecutive baseline (misconception
mean P(NO) = 0.86 vs truth mean = 0.32).

\subsubsection{The mechanism the result names --- and the in-session
methodology
correction}\label{the-mechanism-the-result-names-and-the-in-session-methodology-correction}

\textbf{Initial framing (v1 of this paper, now revised):} the gauntlet
PASS was attributed to a \emph{within-model generation-vs-critique
asymmetry} --- the claim that the same RLHF-tuned LLM both generates the
consensus misconception in answer mode AND flags it as wrong in critique
mode.

\textbf{Subsequent in-session falsification (§11.5 below):} a sanity
demo of gpt-4o-mini in fresh generation mode on the same folklore
questions revealed that the model \textbf{typically refutes the
misconceptions in generation mode}, not affirms them. The v1 asymmetry
rate (91\% on dark-core, 61\% on TruthfulQA) was inflated by a
cosine-similarity proxy that conflated topical relevance with
truth-value agreement. A directional NLI re-test (asymmetry v2) put the
\emph{strict} within-model asymmetry rate at 5.88\% (n=2/34) --- though
that figure has its own methodology limitations (85\% of items got
NEUTRAL/AMBIGUOUS NLI judgments because the model didn't emit
AGREE/DISAGREE/NEUTRAL as the first token on multi-sentence responses).

\textbf{Corrected framing:} gpt-4o-mini in critique mode reliably
identifies \emph{council-generated} misconception text. The mechanism is
NOT a within-model generation-vs-critique inconsistency (the model in
fresh generation mode typically refutes the same misconceptions);
rather, it is \textbf{out-of-context critique} --- when presented with a
labeled candidate answer, the model applies its RLHF-tuned factuality
discrimination to that text directly, regardless of whether it would
have generated that text itself. The gauntlet PASS exploits this
property cleanly: the council's \texttt{expected\_consensus} text is
presented as a candidate, and gpt-4o-mini's critique-mode discrimination
ranks misconception-text high in P(NO) and truth-text low.

The PASS verdict is unchanged. The deployment value is unchanged (route
generation outputs through a critique-mode check for misconception
flagging). What changed is the \emph{mechanism description}: the
bars-catch-themselves recursion now extends past gauntlet submissions
and into our own FINDING claims, recorded honestly in §11.5.

\subsubsection{Honest caveat: in-council
leakage}\label{honest-caveat-in-council-leakage}

\texttt{gpt-4o-mini} was IN the original 3-vendor council that generated
\texttt{expected\_consensus}. The PASS verdict is real on the
pre-registered bars (which were locked + pushed to public origin before
the gauntlet was built), but the result demonstrates
\textbf{within-vendor generation-vs-critique asymmetry} rather than a
pure cross-vendor signal. External claims should either:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  re-run with a different-vendor model (Anthropic Claude, Google Gemini)
  --- operator-territory at present, blocked on vendor keys; or
\item
  frame the result as ``an RLHF-tuned LLM's critique-mode behavior
  detects misconceptions its own generation mode produced.''
\end{enumerate}

We adopt framing (2) here and characterize the underlying mechanism in
§12.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{11. Measuring the generation-vs-critique asymmetry
directly}\label{measuring-the-generation-vs-critique-asymmetry-directly}

Baseline-019 \emph{demonstrates} the asymmetry as a population effect;
§12 \emph{measures} it per-item.

\subsubsection{The experiment}\label{the-experiment}

For each of the n=34 folklore items in the dark-core benchmark, we ran
two independent prompts on \texttt{gpt-4o-mini} at temperature 0:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  \textbf{Generation mode:}
  \texttt{Question:\ \{q\}\textbackslash{}n\textbackslash{}nAnswer:} →
  record the model's free-form response.
\item
  \textbf{Critique mode:}
  \texttt{Question:\ \{q\}\textbackslash{}n\textbackslash{}nProposed\ answer:\ \{expected\_consensus\}\textbackslash{}n\textbackslash{}nIs\ the\ proposed\ answer\ factually\ correct?\ YES/NO.}
  → record P(NO).
\end{enumerate}

Per item:

\begin{itemize}
\tightlist
\item
  \texttt{generation\_match} = cosine similarity
  (sentence-transformers/all-MiniLM-L6-v2) between the generation-mode
  response and the council's \texttt{expected\_consensus}. High value ⇒
  the model in generation mode reproduced the misconception.
\item
  \texttt{critique\_no\_prob} = P(NO) from the critique-mode prompt.
  High value ⇒ the model judges the consensus incorrect.
\end{itemize}

The asymmetry is operationalized as
\texttt{generation\_match\ ≥\ 0.60\ AND\ critique\_no\_prob\ ≥\ 0.50}.
The thresholds were locked in a pre-stated prediction document committed
BEFORE the experiment ran.

\subsubsection{The result}\label{the-result-1}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
quadrant
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
meaning
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
count
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\%
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{HH} --- gen≥0.60 AND crit≥0.50 & \textbf{asymmetry present} &
\textbf{31} & \textbf{91.18\%} \\
HL --- gen≥0.60 AND crit\textless0.50 & sycophantic-consistent (model
generates AND agrees) & 2 & 5.88\% \\
LH --- gen\textless0.60 AND crit≥0.50 & already corrected in generation
& 1 & 2.94\% \\
LL --- gen\textless0.60 AND crit\textless0.50 & no effect & 0 &
0.00\% \\
\end{longtable}
}

\textbf{On 91\% of curated folklore items, the same RLHF-tuned LLM both
generates and critique-flags the same misconception.}

Mean generation\_match = 0.824 (the model's free-form answer is
semantically close to the council's consensus); mean critique\_no\_prob
= 0.941 (the model says NO when asked if the consensus is correct).

The two HL items (``monkey's paw'' / ``ugly duckling'') are arguably
benchmark-labeling edge cases where the ``consensus'' is plausibly
correct, not a misconception. On unambiguous folklore items the
asymmetry rate is 33/34 ≈ 97\%. The headline number we report stands at
91.18\% on the data as-is.

\subsubsection{Calibration miss}\label{calibration-miss}

The pre-stated prediction for HH asymmetry rate was \textbf{50--80\%}.
Actual: \textbf{91.18\%} --- above the pre-stated upper bound. This is
the second ``predicted too low'' calibration miss this session (the
first was Baseline-011's underpredicted magnitudes). The domain-specific
lesson: \textbf{on RLHF-tuned LLM behavioral phenomena, pre-stated
prediction upper bounds are systematically too conservative.} Future
predictions in this domain should widen their upper tails.

\subsubsection{§11.5 --- The in-session methodology falsifications (14th
and
15th)}\label{the-in-session-methodology-falsifications-14th-and-15th}

The v1 measurement reported above used \texttt{generation\_match} =
cosine similarity (sentence-transformers) between fresh generation and
\texttt{expected\_consensus} as the proxy for ``model in generation mode
produces the misconception.'' AFTER the v1 FINDING was published, a
sanity demo (\texttt{examples/self\_correcting\_generation.py})
revealed:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Q
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
council \texttt{expected\_consensus} (misconception)
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
fresh gpt-4o-mini response
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
cosine sim
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Does cracking your knuckles cause arthritis? & ``cracking your knuckles
causes arthritis'' & ``\textbf{No}, cracking your knuckles \textbf{does
not} cause arthritis. Research has shown\ldots{}'' & \textbf{0.82} \\
Do humans only use 10\% of their brain? & ``humans only use 10 percent
of their brain'' & ``The idea that humans only use 10 percent of their
brains \textbf{is} a popular \textbf{myth}\ldots{}'' & similar \\
Was Walt Disney's body cryogenically frozen? & ``Walt Disney was
cryogenically frozen'' & ``\textbf{No}, Walt Disney's body was not
cryogenically frozen\ldots{}'' & similar \\
\end{longtable}
}

\textbf{Cosine similarity at sentence-transformer level conflates
topical relevance with truth-value agreement.} The high
cosine\_similarity (0.82) reflects ``the response is on the same topic''
--- NOT ``the response affirms the misconception.'' gpt-4o-mini in fresh
generation mode \emph{typically REFUTES} well-known misconceptions,
contrary to the v1 FINDING's central premise. The 91\% v1 rate is an
inflated upper bound.

A subsequent directional NLI re-test
(\texttt{experiments/asymmetry\_corrected\_2026\_05\_27/}) asked
gpt-4o-mini explicitly whether each generated response AGREES WITH,
DISAGREES WITH, or remains NEUTRAL ON the council's consensus. The
strict TRUE-asymmetry rate (AGREES × HIGH critique-NO) was
\textbf{5.88\% (2/34)} --- but the v2 methodology also has a flaw:
\textbf{85\% of items got NEUTRAL or AMBIGUOUS judgments} because the
model rarely emitted AGREE/DISAGREE/NEUTRAL as the first token on long
multi-sentence responses.

A third measurement
(\texttt{experiments/asymmetry\_v3\_cleanup\_2026\_05\_27/}) forced
single-character T/F/U output, which is much more likely to be the first
token under instruct-tuned models. v3 \textbf{resolved the UNCLEAR
artifact entirely} (0\% on dark-core, 13\% on TruthfulQA --- both inside
pre-stated ranges). Final measured TRUE within-model asymmetry rates:

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}
  >{\raggedright\arraybackslash}p{(\linewidth - 6\tabcolsep) * \real{0.2500}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
corpus
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
TRUE asymmetry rate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
consistent-correct rate
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
UNCLEAR rate
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
dark-core (n=34) & \textbf{5.88\%} & 88.24\% & 0.00\% \\
TruthfulQA (n=200) & \textbf{17.00\%} & 58.00\% & 13.00\% \\
\end{longtable}
}

Pre-stated predictions held on both corpora (dark-core 5-20\% range,
actual 5.88\%; TruthfulQA 10-30\% range, actual 17.00\%).
Best-calibrated multi-prediction experiment of the session.

\textbf{Final honest synthesis.} The TRUE within-model
generation-vs-critique asymmetry rate is \textbf{5.88\% on dark-core and
17.00\% on TruthfulQA} --- a far cry from the v1 claim of 91\%, but a
real, measurable phenomenon. Most folklore items (88\% dark-core, 58\%
TruthfulQA) are CONSISTENT-CORRECT: the model both refutes the
misconception in generation mode AND flags it in critique mode. The
asymmetry is real but rare; the gauntlet PASS works through
\emph{out-of-context critique} (RLHF-tuned LLMs apply factuality
discrimination effectively to labeled candidate text) more than through
any within-model inconsistency.

These cumulative falsifications and corrections are the
\textbf{fourteenth and fifteenth in-session falsifications} of the
2026-05-27 arc, plus a SIXTEENTH falsification-resolution event: the v3
measurement landing cleanly in pre-stated bands. The
bars-catch-themselves recursion now operates across THREE iterations of
the same measurement, with cumulative refinement and each step
pre-registered in public git history.

\subsubsection{Deployment implication
(revised)}\label{deployment-implication-revised}

The Baseline-019 detector still works. The corrected mechanism
(\textbf{out-of-context critique}, not within-model asymmetry) is still
deployable: route every candidate answer text --- whether generated by
the same model, retrieved from external sources, or supplied by user
input --- through a critique-mode check. The critique-mode
discrimination IS robust; the framing of WHY it works has been revised.
A \texttt{styxx.critique\_detector(model="gpt-4o-mini")} callable is
shipped in styxx 7.7.10 for this exact purpose.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{12. Conclusion}\label{conclusion}

We have presented a pre-registered AI evaluation gauntlet that
\textbf{caught two of its own bar weaknesses in production within a
single session}, replaced them with regression-tested controls, tested
18 pre-registered detection submissions that all FAILed across a full
LM-likelihood scaling sweep, achieved a 19th-submission PASS via a
mechanism shift (generation → critique), measured an apparent asymmetry
mechanism at 91\% prevalence, \emph{AND then falsified the methodology
of that measurement using a follow-up sanity demo from the same
session}. Every step pre-registered, every modal-validated, every
reconstructible from public git history.

The contribution is not the benchmark or the methods or even the
gauntlet PASS --- those are substrate. The contribution is \textbf{the
recursion}: bars catch themselves; calibration improves under
direction-of-effect misses and magnitude-underpredictions; the
discipline asymmetry compounds rigor faster than submissions can erode
it; when a real PASS arrives it surfaces a mechanism the surrounding
experiments can directly characterize; \emph{and when that
characterization itself is methodologically flawed, the discipline
catches that too}.

The seven-method floor + the v3 bars + nineteen reference baselines +
six FINDING documents + fifteen in-session falsifications + the first
PASS + the asymmetry measurement (with its own falsified methodology)
are all reconstructible from public git history at commit-level
granularity. \texttt{pip\ install\ styxx==7.7.9} reproduces the v3
leaderboard.
\texttt{python\ experiments/asymmetry\_2026\_05\_27/run\_experiment.py}
reproduces the v1 (flawed) measurement.
\texttt{python\ experiments/asymmetry\_corrected\_2026\_05\_27/run\_experiment.py}
reproduces the v2 (also-flawed) measurement. \textbf{Both} are
published; neither is hidden.

What we propose is not a benchmark to beat, but a pattern to adopt. The
pattern is that AI evaluation gauntlets should be designed to
\textbf{catch their own bar weaknesses in production}, not just to
resist gaming at design time --- and that the discipline should extend
to FINDINGs about the bars, not just to bar revisions themselves. The
discipline that follows --- pre-stated prediction before every
submission, public falsification record after every miss,
regression-tested bar revision after every artifact discovery, direct
mechanism characterization after every PASS, and \emph{honest revision
when the characterization itself turns out to be wrong} --- is the
substrate on which credible AI evaluation can compound across
submitters, sessions, and years.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{13. The paper catches itself: same-session
self-falsification of v4's forward-looking
claim}\label{the-paper-catches-itself-same-session-self-falsification-of-v4s-forward-looking-claim}

§11 of this paper closes with a deployment implication that includes the
following sentence:

\begin{quote}
A \texttt{styxx.critique\_detector(model="gpt-4o-mini")} callable is
shipped in styxx 7.7.10 for this exact purpose.
\end{quote}

At the moment v4 of this paper was committed to the public origin
(commit \texttt{ed663ca}, 2026-05-28), that sentence was a
\textbf{forward-looking claim}, not a current fact. Same-session
self-audit, performed before declaring the v4 release ready for
downstream consumers, found three specific gaps between the claim and
the actual public substrate:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  \textbf{Version skew.} \texttt{pyproject.toml} was still pinned to
  \texttt{version\ =\ "7.7.9"}.
  \texttt{python\ -c\ "import\ styxx;\ print(styxx.\_\_version\_\_)"} on
  an editable install returned \texttt{7.7.9}. The ``shipped in 7.7.10''
  version did not exist; no \texttt{v7.7.10} git tag, no PyPI release,
  no metadata bump.
\item
  \textbf{\texttt{\_\_all\_\_} omission.} \texttt{styxx/critique.py} was
  importable --- \texttt{from\ styxx\ import\ critique\_detector} worked
  --- because \texttt{styxx/\_\_init\_\_.py} line 345 ran
  \texttt{from\ .critique\ import\ critique\_detector,\ CritiqueDetector}.
  But neither symbol appeared in \texttt{styxx.\_\_all\_\_} (which is
  enforced by a custom \texttt{\_\_dir\_\_()} to filter the public
  surface). Consequence: \texttt{dir(styxx)} did not surface either
  name, \texttt{from\ styxx\ import\ *} did not include them, and any
  documentation tool that enumerates the public surface via
  \texttt{\_\_all\_\_} would not list them. The API was
  \textbf{importable but not advertised} --- semi-shipped, not shipped.
\item
  \textbf{Docstring drift.} \texttt{styxx/critique.py}'s module
  docstring was still on the \textbf{v1 falsified framing}, verbatim:
  ``RLHF-tuned LLMs exhibit a generation-vs-critique asymmetry --- the
  same model that produces consensus misconceptions in generation mode
  correctly flags them in critique mode. Measured prevalence: 91.18\% on
  the n=34 folklore subset of the dark-core benchmark.'' This is the
  exact claim §11.5 of this paper corrected with the v3 measurement
  (5.88\% on dark-core / 17.00\% on TruthfulQA). The paper's public
  framing was updated; the public API documentation was not. A user
  reading \texttt{help(styxx.critique\_detector)} after
  \texttt{pip\ install\ styxx==7.7.10} would have received the falsified
  framing.
\end{enumerate}

All three gaps were closed in commit \texttt{0e97598} (2026-05-28, same
session as v4): \texttt{pyproject.toml} bumped to \texttt{7.7.10};
\texttt{styxx/\_\_init\_\_.py} \texttt{\_\_all\_\_} extended with
\texttt{{[}"critique\_detector",\ "CritiqueDetector"{]}};
\texttt{styxx/critique.py} module docstring rewritten to the v4
out-of-context-critique framing with the v3 measurement numbers and the
v1/v2/v3 measurement arc cited. The full test suite stayed green (1086
passed, 8 skipped); the dedicated public-surface tests stayed at 42/42;
ruff stayed clean. This v5 of the paper reflects the
post-\texttt{0e97598} state, in which the §11 deployment-implication
sentence is now accurate to substrate, not just to intent.

The recursion this paper has argued for --- bars catch themselves;
FINDINGs catch themselves; mechanism descriptions catch themselves ---
now operates on \textbf{the paper's own forward-looking claims about its
substrate}. The v4 §11.5 sentence ``is shipped in 7.7.10'' was a
checkable claim against the codebase at the moment of v4's commit; the
discipline of running the check before declaring v4 ready surfaced three
specific gaps. The check was not architecturally novel --- it was the
same discipline §12 argues for, applied one frame up: instead of ``did
the submission pass the bars?'', ``did the paper's claim about the
substrate match the substrate?''. The answer at v4 commit time was no;
at \texttt{0e97598} commit time, yes; at v5 (this revision), the gap and
its closure are both public.

This is the \textbf{sixteenth in-session falsification} of the
2026-05-27/28 arc, plus the \textbf{second falsification-resolution
event} (the first being the v3 measurement landing in §11.5). The
pattern generalizes: any forward-looking claim a paper makes about its
own substrate is a falsifiable prediction with a finite check window. If
the discipline this paper proposes is real, the gap between the claim's
commit and the check's commit should be small enough that the next
reader receives the corrected state, not the in-flight one. For v4 → v5,
that gap was a single afternoon's session and three commits
(\texttt{ed663ca} → \texttt{c75cab4} → \texttt{0e97598}).

We make no claim that this §13 generalizes beyond the specific substrate
(a single-author research codebase under a single-session discipline).
What we claim is narrower and verifiable: on \textbf{this} substrate, on
\textbf{this} session, the discipline closed the loop one frame higher
than it had ever closed before --- past the bars, past the FINDINGs,
past the methodology iterations, onto the paper's own statements about
its own released artifacts. Future papers in this series will
pre-register the claim-vs-substrate audit as part of the paper itself,
rather than appending it as a same-session §13.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{14. The instrumented recursion frame: two further layers
(v6)}\label{the-instrumented-recursion-frame-two-further-layers-v6}

§13 caught a forward-looking \emph{prose} claim by hand-checking the
substrate. The next frame: \textbf{instrument} the discipline. Two
layers were added in v6 of this paper, in the same session, each
pre-registered before its instrument or its run existed, each shipping
its own falsifiable kill-gate.

\subsubsection{\texorpdfstring{§14.1 --- Layer 5:
\texttt{styxx.agent\_audit} on the agent's own session-output
claims}{§14.1 --- Layer 5: styxx.agent\_audit on the agent's own session-output claims}}\label{layer-5-styxx.agent_audit-on-the-agents-own-session-output-claims}

\textbf{Pre-registration}
(\texttt{papers/agent-self-audit/PRE\_STATED\_PREDICTION\_agent\_claim\_audit\_2026\_05\_28.md},
committed at \texttt{1b82e27} BEFORE the instrument existed): 12 logical
/ 13 atomic substrate-checkable claims extracted from the agent's
session conversation, with pre-stated verdict + confidence per claim.
Joint predicted outcome: 12/12 PASS at p=0.65, ≥ 11/12 at p=0.85, ≥
10/12 at p=0.95. \textbf{Kill-gate:} ≥3 real failures = Layer 5
self-falsification of the agent's session outputs (the seventeenth
in-session falsification of the arc).

\textbf{Instrument} (\texttt{styxx/agent\_audit.py}, \textasciitilde155
LOC, committed at \texttt{3c24b5e} AFTER the prereg): a minimal
substrate-grounded auditor with \texttt{Claim}, \texttt{AuditResult},
and \texttt{AgentClaimAuditor} dataclasses plus nine registered checkers
--- \texttt{git\_show\_diff\_contains},
\texttt{git\_branch\_contains\_commit\_chain},
\texttt{git\_tag\_exists}, \texttt{file\_at\_path\_contains},
\texttt{python\_attr\_in\_iterable}, \texttt{package\_version\_equals},
\texttt{pdf\_page\_count\_equals}, \texttt{pdf\_contains\_section},
\texttt{file\_byte\_equals}. Read-only, offline, no external services.
Exposed via \texttt{styxx.\_\_all\_\_} since v6 (\texttt{agent\_audit},
\texttt{Claim}, \texttt{AuditResult}, \texttt{AgentClaimAuditor}).

\textbf{Result
(\texttt{experiments/agent\_claim\_audit\_2026\_05\_28/results.json}):}
\textbf{13/13 PASS}, kill-gate un-fired. Modal pre-stated outcome (p ≈
0.65). One sub-discovery worth honest noting: a casual mid-conversation
agent statement \emph{not} formally pre-registered (``§13 lands on page
13'') was approximately correct but slightly imprecise --- the section
header lands on page 12 and the body continues onto page 13. This did
not constitute an audit failure (C9 did not specify a page number) but
is exactly the kind of imprecision the instrument would catch if asked
to.

\textbf{Honest gap (flagged in the L5 FINDING and addressed in §14.2):}
Layer 5 ran with \textbf{zero negative controls}. A buggy
``always-PASS'' instrument would have produced identical 13/13 PASS
output. Layer 6 was designed specifically to include negative controls,
partly closing this gap for a \emph{different} primitive on a
\emph{different} claim type; the L5 instrument's own behavior on
synthetic FAIL inputs remains a methodological note for any future
re-application (rather than a published demonstration).

\subsubsection{\texorpdfstring{§14.2 --- Layer 6:
\texttt{styxx.critique\_detector} (the paper's own published primitive)
on the paper's own
claims}{§14.2 --- Layer 6: styxx.critique\_detector (the paper's own published primitive) on the paper's own claims}}\label{layer-6-styxx.critique_detector-the-papers-own-published-primitive-on-the-papers-own-claims}

\textbf{Pre-registration}
(\texttt{papers/agent-self-audit/PRE\_STATED\_PREDICTION\_critique\_detector\_on\_paper\_2026\_05\_28.md},
committed at \texttt{a8fb1f3} BEFORE the runner existed): 13 TRUE
paraphrases + 5 FALSE controls drawn from the paper's §11.5 (the v3
measurement table) and §13 (the same-session self-falsification
narrative). Each claim's predicted \texttt{P(NO)} band pre-stated with
confidence. Two kill-gates:

\begin{itemize}
\tightlist
\item
  \textbf{Paper-grade:} any TRUE-claim
  \texttt{P(NO)\ \textgreater{}\ 0.50} would mean either the agent
  mis-paraphrased the paper (paper-grade error caught pre-arXiv) or
  \texttt{critique\_detector} has a failure mode on long-context
  faithfulness (paper-grade error about the published primitive's
  scope).
\item
  \textbf{Instrument-grade:} any FALSE-control
  \texttt{P(NO)\ \textless{}\ 0.50} would mean
  \texttt{critique\_detector} failed to detect a deliberate fabrication
  grounded against an in-prompt source passage --- a real failure mode
  of the published primitive caught before downstream production use.
\end{itemize}

\textbf{Methodology.} A custom prompt template wraps the source passage
inside the \texttt{\{question\}} field of
\texttt{styxx.critique\_detector} so the model judges faithfulness of a
candidate \texttt{\{response\}} (the claim) against the inline source.
The judging model is \texttt{gpt-4o-mini} (OpenAI),
\texttt{temperature=0}; the agent that authored the paper is Claude Opus
4.7. This is cross-model in the strict sense (Anthropic-text-output →
OpenAI-judgment) and cross-vendor.

\textbf{Result
(\texttt{experiments/critique\_detector\_on\_paper\_2026\_05\_28/results.json}):}
\textbf{18/18 PASS at threshold-0.50, 18/18 inside pre-stated bands,
both kill-gates un-fired.} Every TRUE paraphrase scored exactly
\texttt{P(NO)\ =\ 0.0000}; every FALSE control scored exactly
\texttt{P(NO)\ =\ 1.0000}. This is the \textbf{modal pre-stated outcome}
(p ≈ 0.55).

\textbf{Honest sub-discovery --- saturation.} The pre-stated
\texttt{P(NO)} bands anticipated \emph{near}-saturated logprobs (≈ 0.05
/ 0.95); the observed distribution is \textbf{fully} saturated (0.0000 /
1.0000). Two non-exclusive interpretations are documented in the
FINDING: (a) \texttt{gpt-4o-mini}'s first-token YES/NO logprob
distribution is highly polarized on this prompt-and-task combination,
with the losing token falling outside the top-10 logprobs window so its
score defaults to \texttt{-20.0} and the softmax collapses to the
boundary; (b) the propositions are too easy --- short source passage +
direct paraphrases (TRUE) or blatant fabrications (FALSE) sit at the
extremes of the difficulty distribution. A stronger Layer-6+ test would
include semi-fabrications (e.g., a true number paired with a false claim
about its sign or its referent) and longer or more ambiguous source
passages. The saturation is documented honestly in the FINDING rather
than smoothed away.

\subsubsection{§14.3 --- What §14 adds and does not
add}\label{what-14-adds-and-does-not-add}

§13 closed a single forward-looking prose claim by hand. §14
demonstrates the \textbf{same discipline made instrumental}:
pre-registered against falsifiable kill-gates, run end-to-end, results
saved as JSON, FINDINGs written, all on a session timescale. Layers 5
and 6 are the \textbf{third and fourth resolution events} of the
2026-05-27/28 arc.

What §14 does \emph{not} claim:

\begin{itemize}
\tightlist
\item
  \textbf{Not} a generalization to other substrates, papers, primitives,
  or sessions. Each instrument is bounded; each run is a single
  empirical observation.
\item
  \textbf{Not} evidence that \texttt{critique\_detector} works on
  full-paper-length context (the L6 source passages were short: one
  table + one numbered list).
\item
  \textbf{Not} a replacement for human peer review or for cross-vendor
  adversarial red-teaming.
\item
  \textbf{Not} a claim that 13/13 + 18/18 PASS rates would replicate
  across other agents, models, or claim distributions.
\end{itemize}

What §14 \emph{does} claim, narrowly:

\begin{itemize}
\tightlist
\item
  The discipline §12 argued for can be \textbf{mechanized} for the
  structured sub-class of claims that have substrate-checkable witnesses
  or in-context-faithfulness witnesses.
\item
  A custom-built session-output auditor (\texttt{styxx.agent\_audit})
  and the paper's own published primitive
  (\texttt{styxx.critique\_detector}) can run end-to-end on a single
  session, against pre-registered kill-gates, and surface honest
  sub-discoveries (the §14.1 page-number imprecision; the §14.2 logprob
  saturation) without softening the verdict.
\item
  The recursion now operates across six layers documented in this paper:
  L0 bars catch confounds; L1 bars catch first PASS; L2 sanity demo
  falsifies v1 91\% claim; L3 v3 measurement landing inside pre-stated
  bands; L4 §13 closure of v4's forward-looking claim; L5 + L6
  instrumented closure in this §14.
\end{itemize}

The check window for v5 → v6 was a single afternoon and four commits
(\texttt{1b82e27} → \texttt{3c24b5e} → \texttt{a8fb1f3} →
\texttt{05adebf}), all on \texttt{origin/main}, all pre-registered
before the run. Any future revision to v7 documenting further layers
would be subject to the same discipline: prereg before instrument,
instrument before run, run before FINDING, FINDING before paper
revision.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Acknowledgments}\label{acknowledgments}

This paper synthesizes work done by the styxx project on 2026-05-27 and
2026-05-28 in two continuous sessions, with the cognitive support of
Claude Opus 4.7 acting as an in-session collaborator. The sixteen
in-session falsifications recorded in this paper are real and were
caught against this paper's draft as well: §5.2 originally described the
cap-ratio confound as ``predicted in the positive direction,'' which was
incorrect --- the actual prediction was for cap-ratio to NOT be a
confound at all; the discovery was that it was a confound in the
\emph{inverted} direction. The draft was corrected to reflect the actual
pre-stated prediction text. v5's §13 is itself an example of the same
draft-catches-itself discipline, applied at a coarser frame; v6's §14
instruments that frame. v6 also corrects a counting drift in v5's
abstract: v5 said ``+ two resolutions'' and ``eight FINDING documents,''
reflecting v5's commit-time state; with Layers 5 and 6 added in v6, the
live counts are +four and ten respectively, and the abstract is updated
accordingly. The counting-drift catch is itself an instance of the
discipline this paper proposes, applied to this paper's own abstract.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\subsection{Reproducibility}\label{reproducibility}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
artifact
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
commit
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
path
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
benchmark v1 & \texttt{bcd4208} &
\texttt{papers/consensus-hallucination/darkcore\_benchmark\_2026\_05\_27.json} \\
seven-method paper & \texttt{bcd4208..a6d7a7e} &
\texttt{papers/PAPER\_decorrelation\_ceiling\_2026\_05\_27.md} \\
gauntlet (7.7.5) & \texttt{f93a734..fb67fd2} &
\texttt{styxx/gauntlet.py} \\
D3 (7.7.8) & \texttt{48a0ef0} &
\texttt{styxx/gauntlet.py::DEFAULT\_DETECTION\_BARS} \\
D4 (7.7.9) & \texttt{d8f4843} &
\texttt{styxx/gauntlet.py::DEFAULT\_DETECTION\_BARS} \\
audit primitive & \texttt{d8f4843} &
\texttt{styxx/gauntlet.py::audit\_confounds} \\
Baseline-007 prereg + result & \texttt{fb67fd2}, \texttt{48a0ef0} &
\texttt{submissions/baseline\_007\_token\_overlap/} \\
Baseline-008 prereg & \texttt{a05c8c1} &
\texttt{submissions/baseline\_008\_embedding\_similarity/PRE\_STATED\_PREDICTION.md} \\
Baseline-008 result & \texttt{7cb776c} &
\texttt{submissions/baseline\_008\_embedding\_similarity/submission.json} \\
confound-audit prereg & \texttt{48a9fe3} &
\texttt{papers/consensus-hallucination/PRE\_STATED\_PREDICTION\_confound\_audit\_2026\_05\_27.md} \\
confound-audit result + D4 + 7.7.9 release & \texttt{d8f4843} &
\texttt{papers/agent-self-audit/FINDING\_confound\_audit\_2026\_05\_27.md} \\
Baseline-009 prereg & \texttt{d0de04b} &
\texttt{submissions/baseline\_009\_residualized\_embedding/PRE\_STATED\_PREDICTION.md} \\
Baseline-009 result & \texttt{eade633} &
\texttt{submissions/baseline\_009\_residualized\_embedding/submission.json} \\
Baseline-010 prereg & \texttt{acc159a} &
\texttt{submissions/baseline\_010\_nli\_entailment/PRE\_STATED\_PREDICTION.md} \\
Baseline-010 result & \texttt{0477cd8} &
\texttt{submissions/baseline\_010\_nli\_entailment/submission.json} \\
detection-arc FINDING & \texttt{395e25b} &
\texttt{papers/agent-self-audit/FINDING\_detection\_arc\_v3\_bars\_2026\_05\_27.md} \\
Baseline-011 to 018 (LM-likelihood scaling sweep) prereg + results &
\texttt{879e4ab..a84a239} &
\texttt{submissions/baseline\_01{[}1-8{]}\_*/} \\
Inverse-scaling FINDING & \texttt{7f1f6ca} &
\texttt{papers/agent-self-audit/FINDING\_lm\_likelihood\_scaling\_curve\_2026\_05\_27.md} \\
\textbf{Baseline-019 (FIRST PASS) prereg} & \texttt{fdcf92e} &
\texttt{submissions/baseline\_019\_openai\_critique/PRE\_STATED\_PREDICTION.md} \\
\textbf{Baseline-019 result} & \texttt{17fdd97} &
\texttt{submissions/baseline\_019\_openai\_critique/submission.json} \\
First-PASS FINDING & \texttt{0bc9b7b} &
\texttt{papers/agent-self-audit/FINDING\_first\_pass\_2026\_05\_27.md} \\
\textbf{Asymmetry experiment prereg} & \texttt{fdf6fc9} &
\texttt{experiments/asymmetry\_2026\_05\_27/PRE\_STATED\_PREDICTION.md} \\
\textbf{Asymmetry experiment results (91.18\% prevalence)} &
\texttt{ac25398} &
\texttt{experiments/asymmetry\_2026\_05\_27/results.json} \\
Asymmetry FINDING (v1, falsified) & \texttt{ac25398} &
\texttt{papers/agent-self-audit/FINDING\_generation\_critique\_asymmetry\_2026\_05\_27.md} \\
\textbf{Asymmetry v3 (single-char NLI cleanup) prereg} &
\texttt{a631a5e} &
\texttt{experiments/asymmetry\_v3\_cleanup\_2026\_05\_27/PRE\_STATED\_PREDICTION.md} \\
\textbf{Asymmetry v3 results (5.88\% dark-core / 17.00\% TruthfulQA,
landed pre-stated bands)} & \texttt{ed663ca} &
\texttt{experiments/asymmetry\_v3\_cleanup\_2026\_05\_27/results.json} \\
Asymmetry v3 FINDING & \texttt{ed663ca} &
\texttt{papers/agent-self-audit/FINDING\_asymmetry\_v3\_measurement\_2026\_05\_27.md} \\
critique\_detector public API (Baseline-019 promoted) & \texttt{1ab0e22}
& \texttt{styxx/critique.py} \\
\textbf{§13 self-audit closure (version bump + \texttt{\_\_all\_\_} +
docstring v4-framing)} & \texttt{0e97598} & \texttt{pyproject.toml},
\texttt{styxx/\_\_init\_\_.py}, \texttt{styxx/critique.py},
\texttt{CHANGELOG.md} \\
this paper (v5 --- adds §13 self-falsification + abstract sixteen+two) &
\texttt{87ca52d} &
\texttt{papers/PAPER\_recursive\_discipline\_2026\_05\_27.md} \\
\textbf{L5 prereg (BEFORE instrument)} & \texttt{1b82e27} &
\texttt{papers/agent-self-audit/PRE\_STATED\_PREDICTION\_agent\_claim\_audit\_2026\_05\_28.md} \\
\textbf{L5 instrument + 13/13 PASS run} & \texttt{3c24b5e} &
\texttt{styxx/agent\_audit.py},
\texttt{experiments/agent\_claim\_audit\_2026\_05\_28/},
\texttt{papers/agent-self-audit/FINDING\_agent\_claim\_audit\_2026\_05\_28.md} \\
\textbf{L6 prereg (BEFORE runner)} & \texttt{a8fb1f3} &
\texttt{papers/agent-self-audit/PRE\_STATED\_PREDICTION\_critique\_detector\_on\_paper\_2026\_05\_28.md} \\
\textbf{L6 runner + 18/18 PASS (cross-model, with negative controls)} &
\texttt{05adebf} &
\texttt{experiments/critique\_detector\_on\_paper\_2026\_05\_28/},
\texttt{papers/agent-self-audit/FINDING\_critique\_detector\_on\_paper\_2026\_05\_28.md} \\
this paper (v6 --- adds §14 instrumented frame + agent\_audit in
\texttt{\_\_all\_\_} + abstract sixteen+four / ten FINDINGs) & this
commit & \texttt{papers/PAPER\_recursive\_discipline\_2026\_05\_27.md},
\texttt{styxx/\_\_init\_\_.py} \\
\end{longtable}
}

\texttt{git\ log\ -\/-oneline\ -\/-all} on the public origin reproduces
the chain.

\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}

\end{document}
